\documentclass[conference]{IEEEtran}
\pdfminorversion=7

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsthm}
\usepackage{booktabs,tabularx}
\usepackage{cite,url}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{comment}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{colortbl} 
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}

\setlength{\textfloatsep}{6pt plus 1pt minus 2pt}
\setlength{\floatsep}{6pt plus 1pt minus 2pt}
\setlength{\intextsep}{6pt plus 1pt minus 2pt}
\setlength{\abovecaptionskip}{3pt}

\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}

\begin{document}

\title{TRAPS: Risk-Aware One-Sided Admission Indexing for High-Cost Point Queries}
\author{
  Yadi Wen, Yaxuan Wang, Yujie Xu, Yaling Ren, Dan Yu, Rong Du, Yue Fu, Yongle Chen\\
  \IEEEauthorblockA{
Shanxi Key Laboratory of Industrial Internet Security, Taiyuan University of Technology, China\\
Emails: \{2023005322, 2023005749, 2024002194, 2023006489\}@link.tyut.edu.cn,\\ \{yudan, durong, fuyue, chenyongle\}@tyut.edu.cn
  }
}
\maketitle

\begin{abstract}
We study admission indexing for high-cost point queries. Password verification is our case study: a frontend should reject definite invalid credential pairs before they invoke Argon2id/PBKDF2, while never dropping represented valid credentials. TRAPS separates routing from correctness. A stable learning and risk oracle assigns attempts to regions, and each region is backed by a keyed approximate-membership substrate or compatible keyed tag fingerprint; the backend remains the only accepting authority. TRAPS uses a capped water-filling allocator for regional residual-forward budgets, records backend-confirmed false positives in an exact replay cache, and preserves one-sidedness through stable features and versioned migration. Seeded controlled frontiers, shift grids, dynamic fallback, LANL-conditioned replay, keyed-tag baselines, scale/update, and bounded-worker queueing show the trade-off: keyed tag fingerprints give strong uniform collision bounds, while AMQ-backed TRAPS provides risk-aware residual-channel allocation without placing learning on the authentication-correctness path.
%\keywords{Password authentication, Learned Bloom filter, Credential-flooding defense, Application-layer denial-of-service, Adaptive false-positive control, Capacity-aware pre-screening}

\end{abstract}

\input{context/Introduction}
\input{context/Building Blocks}
\input{context/System Framework and Threat Model}
\input{context/Firewall Construction}
\input{context/Experiments}
\input{context/Related Work}
\input{context/Conclusion}

\section*{AI-Generated Content Acknowledgement}
The authors used OpenAI ChatGPT/Codex to assist with language editing, source-file inspection, reproducibility-script development, artifact packaging, and code-review suggestions. The authors reviewed and verified the resulting manuscript, experiments, code, figures, and claims, and remain responsible for the paper, artifact, and presentation.
\bibliographystyle{IEEEtran}
\bibliography{references}
\end{document}
